Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-13 paper/paper.tex
No vlined
minor changes.

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

@@

\author{Anonymous \and Anonymous\thanks{This work was supported by an anonymous funding agency
[grant numbers 12345 and 67890]} \and Anonymous\samethanks \\
Top Secret Research Lab, Atlantis \\
anonymous@anonymous}

@@

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. The problem arises in many areas, including bioinformatics
\citep{Bonnici:2013}, computer vision \citep{cviu11,pr15}, malware detection
\citep{DBLP:conf/dimva/BruschiMM06}, compilers
\citep{DBLP:conf/cgo/MurrayF12,DBLP:conf/cp/BlindellLCS15}, model checking \cite{Sevegnani:2015}, and pattern recognition
\citep{Conte:2004}, which has lead to substantial research into how best to solve the problem in
practice \cite[and
more]{mcgregor79,Ullman,Cordella:2004,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}.
The problem comes in two forms: in the non-induced variant, edges must be mapped to edges, but the
target may have ``extra edges'', whilst in the induced variant, non-edges must be mapped to
non-edges.

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.5]%[every node/.style={circle, inner sep=0.5mm,node
        %distance=10mm,fill=black}, scale=0.2]

        \begin{scope}[xshift=-1cm, yshift=-3cm, color=black]
        \node at (-2, 1) {$P$:};
        \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
        \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
        \node [fill, circle, inner sep=1.5pt] (c) at (0,1) {};
        \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
        \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};

        \path[solid] (a) edge (b) edge (c) edge (d)
            (e) edge (b) edge (c) edge (d);
        \end{scope}

        \begin{scope}[xshift=4cm, yshift=-3cm, color=black]
        \node at (-2, 1) {$T$:};
        \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
        \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
        \node [fill, circle, inner sep=1.5pt] (c) at (1,2) {};
        \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
        \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};
        \node [fill, circle, inner sep=1.5pt] (f) at (1,0) {};
        \node [fill, circle, inner sep=1.5pt] (g) at (2,0) {};
        \node [fill, circle, inner sep=1.5pt] (h) at (2,2) {};

        \path[solid] (a) edge (b) edge (d)
            (e) edge (b) edge (d)
            (d) edge (c) edge (f)
            (c) edge (h)
            (f) edge (g);
        \end{scope}

    \end{tikzpicture}
    \caption{The pattern graph $P$ on the left cannot be found in the target graph $T$, but if
    the central vertex in the diamond is removed, then a subgraph isomorphism exists.}
    \label{figure:subgraphexample}
\end{figure}

@@

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target---we show an example in \cref{figure:subgraphexample}. (This in some
ways resembles the approximate subgraph matching model of \citet{DBLP:conf/cp/ZampelliDD05},
although we allow any vertex to be removed.) We show that if
$k$ is reasonably small (say, between 1 and 5), then weakened forms of degree and path based
filterings are still effective in pruning the initial search space and in providing additional
constraints respectively. We then show that combining these techniques leads to a practical
algorithm which can scale to work with the families of graphs commonly used to benchmark subgraph
isomorphism algorithms: depending upon the benchmark family, we can close a substantial portion of
the instances, and in many more cases, we can at least obtain an upper bound. This is a significant
improvement over state of the art maximum common subgraph solvers, which cannot even fit many of
these instances in RAM. Finally, we show that iteratively increasing $k$ gives a competitive
algorithm for the maximum common induced subgraph problem.

\subsection{Definitions and Notations} \label{section:definitions}

A non-induced subgraph isomorphism from a graph $P$ to a graph $T$ is an injective mapping $P
\rightarrowtail T $ which maps adjacent vertices to adjacent vertices. An induced subgraph
isomorphism $P \hookrightarrow T$ additionally maps non-adjacent vertices to non-adjacent vertices.
We define a $k$-less subgraph isomorphism from $P$ to $T$ to be a subgraph isomorphism from all but
$k$ vertices of $P$ to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P
\lessind{k} T$ respectively.

The \emph{complement} of a (simple) graph $P$ is the graph $\shortoverline{P}~$.  Adjacent distinct
vertices in $P$ are non-adjacent in $\shortoverline{P}$ while non-adjacent vertices in $P$ are
adjacent in $\shortoverline{P}$.  The \emph{loop complement} of a graph $P$ is the graph
$\loopcomp{P}$.  The construction of $\loopcomp{P}$ is similar to that of $\bar{P}~$, however
whenever a vertex has a loop in $P$ it does not in $\loopcomp{P}$ and vice versa.  When solving an
induced subgraph isomorphism problem we make use of the loop complement; we care about mapping
vertices with loops to vertices with loops, and vertices without loops to vertices without loops.
The following proposition follows directly from the definitions.

@@

There are three easy ways we might try to extend existing algorithms for this problem.
A non-induced $k$-less subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$, with $k$ extra universally-adjacent vertices added to $T$, and so we could try
solving subgraph isomorphism with a modified target graph. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.
Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving each subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.
Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory \citep{DBLP:conf/cp/McCreeshNPS16},
which is extremely problematic on the instances with which we wish to work.

@@

\begin{figure}[t]
    \centering
    \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \begin{scope}
        \node[fill=black] (p) {};
        \node[fill=black] (b) [above of = p,node distance=5mm] {};
        \node[fill=black] (a) [left of = b, node distance=7mm] {};
        \node[fill=black] (c) [right of = b, node distance=7mm] {};
        \node[fill=black] (d) [above of = b,node distance=5mm] {};

        \node[draw=none,right of = p,node distance=3mm,yshift=-0.5mm] {$p$};
        \node[draw=none,left of = a,node distance=3mm,anchor=base] {$a$};
        \node[draw=none,left of = b,node distance=1.8mm,anchor=base] {$b$};
        \node[draw=none,right of = c,node distance=2mm,anchor=base] {$c$};
        \node[draw=none,right of = d,node distance=3mm,anchor=base] {$d$};

        \node[draw=none, left of = a, node distance=8mm] {$P$:};

        \path (p) edge (a) edge (b) edge (c)
            (a) edge (d)
            (b) edge (c) edge (d)
            (c) edge (d);

        \node[fill=gray!50] (a1) [right of = d, node distance = 2.2cm] {};
        \node[fill=black] (b1) [right of = a1, node distance=7mm] {};
        \node[fill=gray!50] (c1) [right of = b1, node distance=7mm] {};
        \node[fill=black] (a2) [below of = a1, node distance=5mm] {};
        \node[fill=black] (b2) [right of = a2, node distance=7mm] {};
        \node[fill=black] (c2) [right of = b2, node distance=7mm] {};
        \node[fill=gray!50] (a3) [below of = a2, node distance=5mm] {};
        \node[fill=black] (b3) [right of = a3, node distance=7mm] {};
        \node[fill=gray!50] (c3) [right of = b3, node distance=7mm] {};

@@

        \node[draw=none, left of = a2, node distance=6mm] {$T$:};

        \path[color=gray!50] (a1) edge (b1) edge (a2)
            (b1) edge (c1)
            (c1) edge (c2)
            (a2) edge (b2) edge (a3)
            (c2) edge (c3)
            (a3) edge (b3)
            (b3) edge (c3);
        \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
            (a2) edge (b3)
            (b2) edge (c2) edge (b3)
            (c2) edge (b3);

        \node[draw=none] (b) [right of = c2, node distance=2cm] {};
        \node[fill=black] (p) [below of = b, node distance=5mm] {};
        \node[fill=black] (a) [left of = b, node distance=7mm] {};
        \node[fill=black] (c) [right of = b, node distance=7mm] {};
        \node[fill=black] (d) [above of = b,node distance=5mm] {};

        \node[draw=none, left of = a, node distance=5mm] {$U$:};

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
        \end{scope}
    \end{tikzpicture}
    \caption{There is a non-induced isomorphism $P\rightarrowtail T$ from the first graph to the
    second, using the highlighted vertices. No such isomorphism exists to the third graph, but there
    is a $1$-less isomorphism $P\lessnonind{1} U$ omitting vertex $b$.}\label{figure:klessexample}
\end{figure}

@@

A vertex $q$ which is adjacent to $p$ is called its \emph{neighbour}.  Let $\N(p) = \{ q \in \V(P) :
 p~\textnormal{and}~q~\textnormal{are adjacent}\}$ be the set of neighbours of $p$.  The \emph{neighbourhood
degree sequence} (NDS) of $p$, $\nds(p)$, is the (non-ascending) sequence of degrees of its
neighbours.

@@

\begin{example}
Consider the pattern graph $P$ and the two target graphs $T$ and $U$ shown in
\cref{figure:klessexample}. The neighbourhood degree sequence of pattern vertex $p$ is
$\nds(p)=(3,3,2)$.
We highlight a subgraph in $T$ which is non-induced isomorphic to $P$. The vertex $p$ can be
mapped to $q$ in the target graph $T$, as $\nds(q)=(5,5,4,2,2)$.  There is a non-induced
mapping of the pattern graph $P$ into $U$ by removing $b$ of $P$.  This removal
changes the neighbourhood degree sequence of $p$ in the $k$-less version of $P$ to $\nds(p)=(2,2)$.
\end{example}

\Cref{figure:klessexample} also illustrates the three cases possible when filtering by neighbourhood
degree sequence.  Removing vertex $d$ causes each entry in $\nds(p)$ to be reduced; removing vertex
$a$ removes an entry from $\nds(p)$; and removing either of vertex $b$ or $c$ in the middle layer
causes both the size of $\nds(p)$ to be reduced and an entry in $\nds(p)$ to be reduced.

@@

As well as reasoning about degrees, we can also reason about paths.  A \emph{path} between two
vertices $p$ and $q$ is a sequence of edges which can be traversed from $p$ to reach $q$. We define
$\paths(p,q,n)$ to be the number of paths of length $n$ between the vertices $p$ and $q$.

>>>>

@@

\author{Anonymous \and Anonymous\thanks{This work was supported by an anonymous funding agency [grant number 12345]} \and Anonymous\thanks{This work was supported by an anonymous funding agency [grant number 67890]} \\
Top Secret Research Lab, Atlantis \\
anonymous@anonymous}

@@

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. This is used in many areas, including bioinformatics \citep{Bonnici:2013},
computer vision \citep{cviu11,pr15}, malware detection \citep{DBLP:conf/dimva/BruschiMM06},
compilers \citep{DBLP:conf/cgo/MurrayF12,DBLP:conf/cp/BlindellLCS15}, and pattern recognition
\citep{Conte:2004}.  The problem comes in two forms: in the non-induced variant, edges must be
mapped to edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges
may only be mapped to non-edges.

@@

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target---we show an example in \cref{fig:subgraphexample}. This in some
ways resembles the approximate subgraph matching model of \citet{DBLP:conf/cp/ZampelliDD05},
although we allow any vertex to be removed. We show that if $k$ is reasonably small (say, between 1
and 5), then weakened forms of degree and path based filterings are still effective in pruning the
initial search space and in providing additional constraints respectively. We then show that
combining these techniques leads to a practical algorithm which can scale to work with the families
of graphs commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark
family, we can close a substantial portion of the instances, and in many more cases, we can at least
obtain an upper bound. This is a significant improvement over state of the art maximum common
subgraph solvers, which cannot even fit many of these instances in RAM. Finally, we show that
iteratively increasing $k$ gives a competitive algorithm for the maximum common induced subgraph
problem.

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=0.6]%[every node/.style={circle, inner sep=0.5mm,node
        %distance=10mm,fill=black}, scale=0.2]

        \begin{scope}[xshift=-1cm, yshift=-3cm, color=black]
        \node at (0, -0.5) {$P$};
        \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
        \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
        \node [fill, circle, inner sep=1.5pt] (c) at (0,1) {};
        \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
        \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};

        \path[solid] (a) edge (b) edge (c) edge (d)
            (e) edge (b) edge (c) edge (d);
        % \node (a2) [below left of = aa,node distance=13mm,xshift=2mm] {};
        % \node (aa2) [above left of = a2,node distance=13mm,xshift=-2mm] {};
        % \node (ab2) [below left of = a2,node distance=13mm,xshift=-2mm] {};
        % \node (ac2) [fill=gray!50,above left of = ab2,node distance=13mm,xshift=-2mm] {};
        % \node (b2) [below right of = aa,node distance=13mm,xshift=-2mm] {};
        % \node (ba2) [above right of = b2,node distance=13mm,xshift=2mm] {};
        % \node (bb2) [below right of = b2,node distance=13mm,xshift=2mm] {};
        % \node (bc2) [fill=gray!50,above right of = bb2,node distance=13mm,xshift=2mm] {};

        % \node (aaa) [below of =aa, node distance=20mm] {};
        %
        % \path[dotted] (a) edge (a1) edge (b1);
        % \path (aa) edge (a1) edge (b1) edge (a2) edge (b2)
        %     (a1) edge (b1)
        %     (aaa) edge (a2) edge (b2)
        %     (a2) edge (aa2) edge (ab2)
        %     (b2) edge (ba2) edge (bb2);
        % \path[color=gray!50] (ac2) edge (ab2) edge (aa2)
        %     (bc2) edge (bb2) edge (ba2);
        %
        % \node [cross] {};
        \end{scope}

        \begin{scope}[xshift=3cm, yshift=-3cm, color=black]
        \node at (0, -0.5) {$T$};
        \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
        \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
        \node [fill, circle, inner sep=1.5pt] (c) at (1,2) {};
        \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
        \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};
        \node [fill, circle, inner sep=1.5pt] (f) at (1,0) {};
        \node [fill, circle, inner sep=1.5pt] (g) at (2,0) {};
        \node [fill, circle, inner sep=1.5pt] (h) at (2,2) {};

        \path[solid] (a) edge (b) edge (d)
            (e) edge (b) edge (d)
            (d) edge (c) edge (f)
            (c) edge (h)
            (f) edge (g);

        % \node (a2) [below left of = aa,node distance=13mm,xshift=2mm] {};
        % \node (aa2) [above left of = a2,node distance=13mm,xshift=-2mm] {};
        % \node (ab2) [below left of = a2,node distance=13mm,xshift=-2mm] {};
        % \node (ac2) [fill=gray!50,above left of = ab2,node distance=13mm,xshift=-2mm] {};
        % \node (b2) [below right of = aa,node distance=13mm,xshift=-2mm] {};
        % \node (ba2) [above right of = b2,node distance=13mm,xshift=2mm] {};
        % \node (bb2) [below right of = b2,node distance=13mm,xshift=2mm] {};
        % \node (bc2) [fill=gray!50,above right of = bb2,node distance=13mm,xshift=2mm] {};

        % \node (aaa) [below of =aa, node distance=20mm] {};
        %
        % \path[dotted] (a) edge (a1) edge (b1);
        % \path (aa) edge (a1) edge (b1) edge (a2) edge (b2)
        %     (a1) edge (b1)
        %     (aaa) edge (a2) edge (b2)
        %     (a2) edge (aa2) edge (ab2)
        %     (b2) edge (ba2) edge (bb2);
        % \path[color=gray!50] (ac2) edge (ab2) edge (aa2)
        %     (bc2) edge (bb2) edge (ba2);
        %
        % \node [cross] {};
        \end{scope}

    \end{tikzpicture}
    \caption{The pattern graph $P$ on the left cannot be found in the target graph $T$, but if
    the central vertex in $P$ is removed, then a subgraph isomorphism exists.}
    \label{fig:subgraphexample}
\end{figure}
\subsection{Definitions and Notations} \label{section:definitions}

A \emph{non-induced subgraph isomorphism} from a graph $P$ to a graph $T$ is an injective mapping $P
\rightarrowtail T $ which maps adjacent vertices to adjacent vertices. An \emph{induced subgraph
isomorphism} $P \hookrightarrow T$ additionally maps non-adjacent vertices to non-adjacent vertices.
We define a \emph{$k$-less subgraph isomorphism} from $P$ to $T$ to be a subgraph isomorphism from all but
$k$ vertices of $P$ to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P
\lessind{k} T$ respectively.

The \emph{complement} of a (simple) graph $P$ is the graph $\shortoverline{P}~$.  Adjacent distinct
vertices in $P$ are non-adjacent in $\shortoverline{P}$ while non-adjacent vertices in $P$ are
adjacent in $\shortoverline{P}$.  The \emph{loop complement} of a graph $P$ is the graph
$\loopcomp{P}$.  The construction of $\loopcomp{P}$ is similar to that of $\shortoverline{P}~$, however
whenever a vertex has a loop in $P$ it does not in $\loopcomp{P}$ and vice versa.  When solving an
induced subgraph isomorphism problem we make use of the loop complement; we care about mapping
vertices with loops to vertices with loops, and vertices without loops to vertices without loops.
The following proposition follows directly from the definitions.

@@

There are three easy ways we might try to extend existing algorithms for this problem.
A non-induced $k$-less subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$, with $k$ extra universally-adjacent vertices added to $T$, and so we could try
solving subgraph isomorphism with a modified target graph. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.
Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving each subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.
Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory \citep{DBLP:conf/cp/McCreeshNPS16},
which is extremely problematic on the instances we wish to work with.

@@

\begin{figure}[t]
    \centering
    \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[fill=black] (b) [above of = p,node distance=5mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=5mm] {};

        \node[draw=none,right of = p,node distance=3mm,yshift=-1mm] {$p$};
        \node[draw=none,left of = a,node distance=3mm,yshift=-1mm] {$a$};
        \node[draw=none,left of = b,node distance=3mm,yshift=-1mm] {$b$};
        \node[draw=none,right of = c,node distance=3mm,yshift= 1mm] {$c$};
        \node[draw=none,above of = d,node distance=3mm,yshift=1mm] {$d$};

        \node[draw=none, below of = p, node distance=6mm] {$P$};

        \path (p) edge (a) edge (b) edge (c)
            (a) edge (d)
            (b) edge (c) edge (d)
            (c) edge (d);
        \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=gray!50] (a1) [right of = d, node distance = 2cm] {};
        \node[fill=black] (b1) [right of = a1] {};
        \node[fill=gray!50] (c1) [right of = b1] {};
        \node[fill=black] (a2) [below of = a1, node distance=6mm] {};
        \node[fill=black] (b2) [right of = a2] {};
        \node[fill=black] (c2) [right of = b2] {};
        \node[fill=gray!50] (a3) [below of = a2, node distance=6mm] {};
        \node[fill=black] (b3) [right of = a3] {};
        \node[fill=gray!50] (c3) [right of = b3] {};

@@

        \node[draw=none, below of = b3, node distance=6mm] {$T$};

        \path[color=gray!50] (a1) edge (b1) edge (a2)
            (b1) edge (c1)
            (c1) edge (c2)
            (a2) edge (b2) edge (a3)
            (c2) edge (c3)
            (a3) edge (b3)
            (b3) edge (c3);
        \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
            (a2) edge (b3)
            (b2) edge (c2) edge (b3)
            (c2) edge (b3);
    \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[draw=none] (b) [above of = p,node distance=5mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=5mm] {};

        \node[draw=none, below of = p, node distance=6mm] {$U$};

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
    \end{tikzpicture}
    \caption{There is a non-induced isomorphism $P\rightarrowtail T$ from the first graph to the
    second, using the highlighted vertices. No such isomorphism exists to the third graph, but there
    is a $1$-less isomorphism $P\lessnonind{1} U$ omitting vertex $b$.}\label{fig:klessexample}
\end{figure}

@@

A vertex $q$ which is adjacent to $p$ is called its \emph{neighbour}.  Let $\N(p) = \{ q \in \V(P) :
 p~\textnormal{and}~q~\textnormal{are adjacent}\}$ be the \emph{set of neighbours} of $p$.  The \emph{neighbourhood
degree sequence} (NDS) of $p$, $\nds(p)$, is the (non-ascending) sequence of degrees of its
neighbours.

@@

\begin{example}
Consider the pattern graph $P$ and the two target graphs $T$ and $U$ shown in
\cref{fig:klessexample}. The neighbourhood degree sequence of pattern vertex $p$ is
$\nds(p)=(3,3,2)$.
We highlight a subgraph in $T$ which is non-induced isomorphic to $P$. The vertex $p$ can be
mapped to $q$ in the target graph $T$, as $\nds(q)=(5,5,4,2,2)$.  There is a non-induced
mapping of the pattern graph $P$ into $U$ by removing $b$ of $P$.  This removal
changes the neighbourhood degree sequence of $p$ in the $k$-less version of $P$ to $\nds(p)=(2,2)$.
\end{example}

\Cref{fig:klessexample} also illustrates the three cases possible when filtering by neighbourhood degree
sequence.  Removing the top node causes each entry in $\nds(p)$ to be reduced; removing the
left-middle node removes an entry from $\nds(p)$; and removing either of the connected nodes in the
middle layer causes both the size of $\nds(p)$ to be reduced and an entry in $\nds(p)$ to be reduced.

@@

As well as reasoning about degrees, we can also reason about paths.  A \emph{path} between two
vertices $p$ and $q$ is a sequence of edges which can be traversed from $p$ to reach $q$. We define
$\paths(p,q,n)$ to be the \emph{number of paths of length $n$} between the vertices $p$ and $q$.
